\documentclass{article}
\usepackage{amsmath}


\newcommand{\field}[1]{\item \texttt{#1}}
\newcommand{\num}[1]{\setcounter{enumi}{#1}}

\begin{document}
\title{Homework 1 - Theory}
\author{Patrick Collins}
\date{\today}
\maketitle

\begin{enumerate}
\setcounter{enumi}{2}

% safety -- bad thing never happens
% liveness -- good things eventually happen
\item Identify safety and liveness properties. 
  \begin{enumerate}
  \item Liveness. The ``good thing'' is being served.
  \item Liveness. The ``good thing'' is receiving the ball that has
    beeen thrown.
  \item Safety. The ``bad thing'' is waiting for longer than it takes
    for the first thread to exit its critical section.
  \item Safety. The ``bad thing'' is waiting more than one second to
    print a message.
  \item Liveness. The ``good thing'' is priting a message.
  \item Safety. The ``bad thing'' is the cost of living decreasing.
  \item Liveness. The ``good thing'' is taxes being paid and resources
    being freed by death.
  \item Safety. The ``bad thing'' is to be identified as something
    other than a Harvard man.
  \end{enumerate}

\item Design a producer-consumer protocol with interrupts. \\
Alice and Bob each have a can on their windowsill. Alice does the following:
\begin{enumerate}
\item She waits until her can is down.
\item She releases the pets.
\item When the pets returns, Alice checks whether they finished the
  food. If they did, she resets her can and knocks over Bob's.\\
\end{enumerate}
Bob does the following:
\begin{enumerate}
\item He waits until his can is down.
\item He puts food in the yard.
\item He pulls the string and knocks Alice's can down.
\end{enumerate}

\item
\item Design a strategy that allows $P - 1$ of $P$ prisoners to be
  freed.\\
The first prisoner determines the parity of blue hats in front of
him. If it is even, he says ``Blue'', otherwise he says ``Red.'' Every
prisoner then determines the parity of blue hats in front of
him. Every prisoner then begins to think ``Even'' or ``Odd.'' Each
prisoner guesses accordingly: if his hat must be blue for the parity
that he is thinking of to be correct, he answers ``Blue,'' otherwise
he answers ``Red.'' Whenever a prisoner hears another answers blue, he
starts to think ``Odd'' if he was thinking ``Even'' before, and
``Even'' otherwise. This strategy allows every prisoner except for the
first to guess the color of his own hat.
% do we need to prove this?

\num{8}

\item 